---
output:
  html_document: default
  pdf_document: default
---
## White Wine Quality by Reem Tarek 
========================================================

##   This dataset is public available for research. The details are described in [Cortez et al., 2009]. 
##  Please include this citation if you plan to use this database:

##  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
##  Modeling wine preferences by data mining from physicochemical properties.
## In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

## Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
##               [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
##                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

# Number of Instances: 4898 records for white wine. 



```{r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
install.packages("knitr",repos="http://cran.us.r-project.org")
install.packages("dplyr",repos="http://cran.us.r-project.org")
install.packages("rmarkdown",repos="http://cran.us.r-project.org")
install.packages('rprojroot', repos="http://cran.us.r-project.org")
install.packages("psych",repos="http://cran.us.r-project.org")

```

```{r}
library(ggplot2)
library(knitr)
library(dplyr)
library(rmarkdown)
library(psych)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
# create a data frame called WW that holds all the white wine data

ww <- read.csv("E:/NanoDegree Data Analyst/EDA/project/wineQualityWhites.csv", sep = ",")

```

> This dataset consists of 4898 obseration of white wine and its quality rating as stated by some of the wine tasters world wide.
> it has 11 variables:
>  Input variables (based on physicochemical tests):
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)
>  No missing values are found.
> Description of attributes:
 1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
 2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
 3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
 4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with  greater than 45 grams/liter are considered sweet
 5 - chlorides: the amount of salt in the wine
 6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
11 - alcohol: the percent alcohol content of the wine
12 - quality (score between 0 and 10)


# Univariate Plots Section


```{r echo=FALSE, Wine_Rating}
names(ww)
ww$rating <- ifelse(ww$quality < 5, 'bad', ifelse(
  ww$quality < 7, 'average', 'good'))
ww$rating <- ordered(ww$rating, levels = c('bad', 'average', 'good'))
ww


```
### Summary:
#### Facotrized the quality into categorical rating for the upcoming analysis

```{r echo=FALSE, Univariate_Plot_Wine_rating}

qplot(rating, data=ww)

```
### Summary:
#### Plotted a Histogram of the derived ratings

```{r echo=FALSE, Univariate_Plot_Wine_quality}


qplot(factor(quality), data=ww)
```
### Summary:
#### Plotted a Histogram of the quality, it is normally distributed with a median of 6.

```{r echo=FALSE, Univariate_Fixed_acidity}
## Fixed acidity in (tartaric acid - g / dm^3)
###Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
###3.800   6.300   6.800   6.855   7.300  14.200 
  
  
qplot(fixed.acidity, data=ww)
summary(ww$fixed.acidity)

```
### Summmay:
#### Fixed Acidity is normally distributed with mean of 6.9

```{r echo=FALSE, Univariate_Plot_Volatile_acidity}
## Volatile acidity (acetic acid - g / dm^3):
### Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
### 0.0800  0.2100  0.2600  0.2782  0.3200  1.1000 

qplot(volatile.acidity, data=ww)
summary(ww$volatile.acidity)
```
### Summary:
#### Volatile Acidity is normally distributed with a mean of 0.28

```{r echo=FALSE, Univariate_Plots_Citric_Acid}

## Citric Acid:
###   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
### 0.0000  0.2700  0.3200  0.3342  0.3900  1.6600 


qplot(citric.acid, data=ww, binwidth=.1)
summary(ww$citric.acid)
```
### Summary:
#### Volatile Acidity is normally distributed with a mean of 0.33


```{r echo=FALSE, Univariate_Plots_Resdidual_Sugar}
##  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 0.600   1.700   5.200   6.391   9.900  65.800 
qplot(residual.sugar, data=ww)

summary(ww$residual.sugar)
```

### Summary:
#### Residual sugar could have been normally distributed but it has a peek values in the frst Quad
```{r echo=FALSE, Univariate_Plots_Chlorides}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 0.00900 0.03600 0.04300 0.04577 0.05000 0.34600 
qplot(chlorides, data=ww)
summary(ww$chlorides)
```
### Summary:
#### Chlorides are normally distributed with mean of 0.046
```{r echo=FALSE, Univariate_Plots_Sulfur_Dioxide}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   2.00   23.00   34.00   35.31   46.00  289.00 
qplot(free.sulfur.dioxide, data=ww)
summary(ww$free.sulfur.dioxide)
```
### Summary:
#### Free Sulfur Dioxide is normally distributed with mean 35.31

```{r echo=FALSE, Univariate_Plots_Total_Sulfur_Dioxide}
##   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##    9.0   108.0   134.0   138.4   167.0   440.0 
qplot(total.sulfur.dioxide, data=ww)

summary(ww$total.sulfur.dioxide)
```
### Summary: 
#### Total SulfurDioxide is normally distributed wih mean of 138.

```{r echo=FALSE, Univariate_Plot_Wine_Desnisty}

##  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 0.9871  0.9917  0.9937  0.9940  0.9961  1.0390 
qplot(density, data=ww)
summary(ww$density)
```
### Summary
#### Density is normally distributed with mean of 0.99

```{r echo=FALSE, Univariate_Plot_Wine_PH_Value}
##  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 2.720   3.090   3.180   3.188   3.280   3.820 
qplot(pH, data=ww)
summary(ww$pH)
```
### Summary
#### PH value is normally distributed with mean values 3.18, note that acid has PH valu < 6 as part of their chemical properties
```{r echo=FALSE, Univariate_Plot_Sulphates}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
## 0.2200  0.4100  0.4700  0.4898  0.5500  1.0800 
qplot(sulphates, data=ww)
summary(ww$sulphates)

```
### Summary
#### Sulphates is almost normally distributed, but it has two peeks. 
```{r echo=FALSE, Univariate_Plot_Alcohol}

###Alcohol:
#### Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
####  8.00    9.50   10.40   10.51   11.40   14.20 

qplot(alcohol, data=ww)
summary(ww$alcohol)
```
### Summary
#### Alcohol is not normally distributed and it has several peeks

# Univariate Analysis


### What is the structure of your dataset?
#### The Dataset consists of almost 5000 records of different white wine samples. it has 11 variables and the quality stated for each one.
#### I have added the column rating based on the quality, the rating includes bad, average and good as a categorical classification of the data.



### What is/are the main feature(s) of interest in your dataset?

####  1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
#### 2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
####  3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
####  4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with  greater than 45 grams/liter are considered sweet
####  5 - chlorides: the amount of salt in the wine
####  6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
#### 7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
#### 8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content
#### 9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
#### 10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
#### 11 - alcohol: the percent alcohol content of the wine
#### 12 - quality (score between 0 and 10)


## What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
#### the rating feature concluded from the quality variabl.
## Did you create any new variables from existing variables in the dataset?
#### yes, the wine rating based on the quality has been dervied.

## Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
#### There were no unusual distibutions among all the variables, all the attributes are normally distibuted except for the alcohol variable, it had several peaks 

---

# Bivariate Plots Section

```{r echo=FALSE,correlation_matrix }
## Correlation between Variables:
 #### by removing the X and rating columns 

correlation_matrix <- cor(ww %>%
    dplyr::select(-X) %>%
    dplyr::select(-rating) %>%
    mutate(
      quality = as.numeric(quality)
    )
  )
## Results:
View(correlation_matrix)
```

### Summary:
#### As part of the bivariate analysis, the correlation between the variables should be calulated to define the strength and direction of the relationships between the variables.

```{r echo=FALSE, Bivariate_Plot_Density_Alcohol}

ggplot(aes(x=alcohol,y=density), data=ww)+
  geom_point(alpha=1/10)

```
### Summary:
#### one of the most interesting bivariate relationships is the alcohol-density one,before plotting it, I expected that as much as the alcohol increases the density decreases and that was proved to be totally correct through the plot and the correltion value:-0.780


```{r echo=FALSE, Bivariate_Plot_Chlorides_Alcohol}

ggplot(aes(x=chlorides,y=alcohol), data=ww)+
  geom_point(alpha=1/10)
```
### Summary:
#### the chlorides and alcohol relationship is strong and the correlation is -0.36018871

```{r echo=FALSE, Bivartiate_plot_Alchol_Quality}
ggplot(aes(x=alcohol,y=quality), data=ww)+
  geom_point(alpha=1/10)+geom_jitter()

```
### Summary
#### prior to the plot, I thought there is no significant relationship between the quality and alcohol,
#### butlooks like, I was wrong, there is a correlation of 0.4 between boht of them.

```{r echo=FALSE, Bivartiate_plot_Density_Quality}
ggplot(aes(x=density,y=quality), data=ww)+
  geom_point(alpha=1/20)+
  geom_jitter()
```

### Summary:
#### Quality and density sort of have a  mild negative relationship with correlation of -0.307123313


```{r echo=FALSE, Bivartiate_plot_Sugar_Density}
ggplot(aes(x=residual.sugar,y=density), data=ww)+
  geom_point(alpha=1/10)
```
### Summary:
#### Prior to the plot, the more you add sugar to a liquid, the higher its density gets, the plot proved this true strong positive correlation 0.83896645


---

# Bivariate Analysis
### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

Regarding the quality, Alcohol, shows a high direct correlation with it with  a value of 0.4.
then comes the density and the quality amount, with high negative correlation 


the highest correlation in general happens to be between:
1- Density and Alcohol, the more you add the alcohol, the less dense the wine becomes (-0.78)
2- Density and Residual Sugar, the more the sugar is added the denser the wine gets (0.83)
3- Also there is a correlation between Total Sulfur Dioxide and Residual Sugar, Density and Free Sulfur Dioxide with the values (0.4, 0.5298, 0.615) respectively 



### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
there is a correlation between Total Sulfur Dioxide and Residual Sugar, Density and Free Sulfur Dioxide with the values (0.4, 0.5298, 0.615) respectively 

### What was the strongest relationship you found?
Density and Residual Sugar, the more the sugar is added the denser the wine gets (0.83)

---
# Multivariate Plots Section


```{r echo=FALSE, Multivariate_Plot_Alcohol_Density_Alcohol}

ggplot(aes(x = alcohol, y = density), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')

```
### Summary:
#### Previously in the bi-variate section, the density the alcohol had a negative relationship, 
#### but it doesn't show how does that affect the quality of the wine. but also it was shown that high  quality wine bottles has lower density
###  and this plot combines the three of them and proves the same

```{r echo=FALSE, Multivariate_Alcohol_Chlorides_Quality}

  
ggplot(aes(x = alcohol, y = chlorides), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')

```
### Summary:
#### From the correlation Matrix, the chorides have a weak negative relationship with the quality, in other words, its seems that best wine bottles have fewer amounts of chlorides. but on the other hand the higher the alcohol amounts are the better the wine is, this plot combines both results, you may notice that the dark spot is at high alcholol/low chlorides quad.
```{r echo=FALSE, Multivariate_Plot_Sugar_Density_Quality}

## Adding the quality to the previous Residual Sugar vs Density plot:
###  Resdidual Sugar and density apprenttly have direct relation with each other but  the quality of the wine depends
ggplot(aes(x = residual.sugar, y = density), data = subset(ww, residual.sugar < 30)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')

```
### Summary:
#### The Resdidual sugar and the density have a strong positive relationship, and it was shown that the better wine types seems to have lower density, this plot brings the three variables toghether, 
##the darker spots of the better quality are shown to be on the lower side of the graph along with the low density.

```{r echo=FALSE, Multivariate_Plot_Sulphates_alcohol_quality}

## Sulphats vs alcohol vs Quality

ggplot(aes(x = alcohol, y = sulphates), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')

```
### Summary:
#### Wines with higher levels of sulphates and alchol are more likely to have higher quality.

```{r echo=FALSE, Multivariate_Plot_Acids_Quality}

## Acids vs Quality

### no Variance noted between the Volatile acids, citric acids vs the quality 
ggplot(aes(x = citric.acid, y = volatile.acidity), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```



```{r echo=FALSE, Multivariate_Plot_Fixed_Acids_Quality}

## Acids vs Quality
### Citric acids vs the fixed acidity.
### no Variance noted between the Volatile acids, citric acids vs the quality 
ggplot(aes(x = citric.acid, y = fixed.acidity), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```

```{r echo=FALSE, Multivariate_Plot_All_Acids_Quality}

## Acids vs Quality
### Volatile acidity vs the fixed acidity.
### no Variance noted between the Volatile acids, citric acids vs the quality 
ggplot(aes(x = volatile.acidity, y = fixed.acidity), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```
### Summary of the previous three plots.

#### the acid levels do not show any variance with the quality as a variable in this data set

```{r echo=FALSE, Multivariate_oxides_quality}

## Sulfur dioxide  vs Quality

### Both oxides are dependent on each other but quality does not depend on any of them.
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```
### Summary:
#### Free sulfur dioxide and total sulfur dioxide have a stron positive correlation of 0.615500965. but seems that the distribution of win quality does not vary between the 3 facets, as if both of them are constants in the wine quality equation.


```{r echo=FALSE, Multivariate_Sulphates_Density_quality}


##Sulphates vs Density  vs Quality

### Sulphates increase the wine quality
ggplot(aes(x = sulphates, y = density), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
  facet_wrap(~rating) 

```
### Summary:
#### From the correlation matrix, sulphates has relativley high correlation coefficient with density, but ironically the values still insignificant (0.07449315), and with comparing both with the quality, it seems there a small variance accross the three facets of the quality in the plot that favorizes the sulphates amounts

---

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

- Wine with higher levels of alcohol appears to be of high quality, regardless of the density and the amount of sugar.
- Wine does not really depend on the acids 


### Were there any interesting or surprising interactions between features?
- Yes, the relationshp between the alcohol amount and the wine quality.

- And of course the higher the resdiual sugar gets the denser the wine, but in all cases quality does not depend on any of them.
- - Wine does not really depend on the acids 

---

### Plot One
```{r echo=FALSE, Alchohol_VS_Quality }

ggplot(aes(x = factor(quality), y = alcohol), data = ww) + geom_boxplot() +
  stat_summary(fun.y = mean, size = 1, color = 'red', geom = 'point') + 
  ylab("Alcohol (percentage of volume)") +
  xlab("Wine Quality (scaled from 0-10)") +
  ggtitle("Alcohol vs wine quality") 
```


### Description One

#### This plot, a box plot, that distributes the alcohol quantity across the factorized values of the quality. the red point describe the mean of the alcohol level at each quality values, and obviously, it in increases directly with the wine quality. 


### Plot Two
```{r echo=FALSE, Chloride_vs_Alcohol}
ggplot(aes(x = chlorides, y = alcohol),data = ww )+ 
  geom_point(aes(color = factor(quality))) +
  labs(x = expression(Chlorides~(g/dm^3)),
       y = "Alcohol (percentage of volume)") + 
  scale_color_brewer(type = 'seq',guide = guide_legend(title = 'Wine Quality', reverse = T,override.aes = list(alpha = 1,size = 2))) +
  ggtitle("Alcohol by Chlorides for Differing Quality White Wines")
```

### Description Two
#### In this plot,the relationship between alcohol and chlorides is drived and how they both relate to quality. Wines with less alcohol, and more chlorides tend to be of low quality.While wines with more alcohol and less chlorides tend to be of low quality. 


### Plot Three
```{r echo=FALSE, Sulphates_vs_Alcohol}
ggplot(aes(x = sulphates, y = alcohol),data = ww )+ 
  geom_point(aes(color = factor(quality))) +
  labs(x = expression(Sulphates~(g/dm^3)),
       y = "Alcohol (percentage of volume)") + 
  scale_color_brewer(type = 'seq',guide = guide_legend(title = 'Wine Quality', reverse = T,override.aes = list(alpha = 1,size = 2))) +
  ggtitle("Alcohol by Sulphates for Differing Quality White Wines")
```

### Description Three
#### High alcohol contents and high sulphate concentrations combined seem to produce better wines.
------
  
# Reflection

At the begining,I started with derving a histogram for each variable in the dataset, all the variables were normally distributed.


in the bivariate analysis, I created a correlation matrix, to check all the coleration values between all the variables. and accordinfly I started plotting the scartter diagram between certain variables and accordingly in the mulitvariate analysis

I think this was a good way to start, and I tried not to spend too much time experimenting with variables that didn't have an obvious correlation. So I was succesful with most obvious correlations, or looking at trends that had the highest correlation coefficient. I had struggles with the variables relating to sulfur dioxide or whenever I tried to create a new variable to reveal latent trends or better visual trends already apparent.  
in the future, I think I might add the red wine dataset to the analysis, expermint the same on it alone and then both dataset againt each other.