White Wine Quality by Reem Tarek 
========================================================

##   This dataset is public available for research. The details are described in [Cortez et al., 2009]. 
##  Please include this citation if you plan to use this database:

##  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
##  Modeling wine preferences by data mining from physicochemical properties.
## In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

## Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016
##               [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf
##                [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib

# Number of Instances: 4898 records for white wine. 


> **Tip**: You will see quoted sections like this throughout the template to
help you construct your report. Make sure that you remove these notes before
you finish and submit your project!

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.



install.packages("ggplot2", dependencies = T) 
install.packages("knitr", dependencies = T)
install.packages("dplyr", dependencies = T)
install.packages("rmarkdown")
install.packages('rprojroot', dependencies=TRUE, repos='http://cran.rstudio.com/')
install.packages("psych")
library(ggplot2)
library(knitr)
library(dplyr)
library(rmarkdown)
library(psych)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
# create a data frame called WW that holds all the white wine data

ww <- read.csv("E:/NanoDegree Data Analyst/EDA/project", sep = ",")
View(ww)

```

> This dataset consists of 4898 obseration of white wine and its quality rating as stated by some of the wine tasters world wide.
> it has 11 variables:
>  Input variables (based on physicochemical tests):
   1 - fixed acidity (tartaric acid - g / dm^3)
   2 - volatile acidity (acetic acid - g / dm^3)
   3 - citric acid (g / dm^3)
   4 - residual sugar (g / dm^3)
   5 - chlorides (sodium chloride - g / dm^3
   6 - free sulfur dioxide (mg / dm^3)
   7 - total sulfur dioxide (mg / dm^3)
   8 - density (g / cm^3)
   9 - pH
   10 - sulphates (potassium sulphate - g / dm3)
   11 - alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)
>  No missing values are found.
> Description of attributes:
 1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
 2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
 3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
 4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with  greater than 45 grams/liter are considered sweet
 5 - chlorides: the amount of salt in the wine
 6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content
9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
11 - alcohol: the percent alcohol content of the wine
12 - quality (score between 0 and 10)


# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
names(ww)
ww$rating <- ifelse(ww$quality < 5, 'bad', ifelse(
  ww$quality < 7, 'average', 'good'))
ww$rating <- ordered(ww$rating, levels = c('bad', 'average', 'good'))
ww
```

```{r echo=FALSE, Univariate_Plots}

qplot(rating, data=ww)
qplot(quality, data=ww)
```
```{r echo=FALSE, Univariate_Plots}
## Fixed acidity in (tartaric acid - g / dm^3)
###Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
###3.800   6.300   6.800   6.855   7.300  14.200 
  
  
qplot(fixed.acidity, data=ww)
summary(ww$fixed.acidity)

```
```{r echo=FALSE, Univariate_Plots}
## Volatile acidity (acetic acid - g / dm^3):
### Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
### 0.0800  0.2100  0.2600  0.2782  0.3200  1.1000 

qplot(volatile.acidity, data=ww)
summary(ww$volatile.acidity)
```

```{r echo=FALSE, Univariate_Plots}

## Citric Acid:
###   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
### 0.0000  0.2700  0.3200  0.3342  0.3900  1.6600 


qplot(citric.acid, data=ww, binwidth=.1)
summary(ww$citric.acid)
```
```{r echo=FALSE, Univariate_Plots}
qplot(residual.sugar, data=ww)
```
```{r echo=FALSE, Univariate_Plots}
qplot(chlorides, data=ww)
```

```{r echo=FALSE, Univariate_Plots}
qplot(free.sulfur.dioxide, data=ww)
```

```{r echo=FALSE, Univariate_Plots}
qplot(total.sulfur.dioxide, data=ww)
```
```{r echo=FALSE, Univariate_Plots}
qplot(density, data=ww)
```

```{r echo=FALSE, Univariate_Plots}
qplot(pH, data=ww)
```

```{r echo=FALSE, Univariate_Plots}
qplot(sulphates, data=ww)
```

```{r echo=FALSE, Univariate_Plots}

###Alcohol:
#### Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
####  8.00    9.50   10.40   10.51   11.40   14.20 

qplot(alcohol, data=ww)
summary(ww$alcohol)
```








> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
#### The Dataset consists of almost 5000 records of different white wine samples. it has 11 variables and the quality stated for each one.
#### I have added the column rating based on the quality, the rating includes bad, average and good as a categorical classification of the data.



### What is/are the main feature(s) of interest in your dataset?

####  1 - fixed acidity: most acids involved with wine or fixed or nonvolatile (do not evaporate readily)
#### 2 - volatile acidity: the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
####  3 - citric acid: found in small quantities, citric acid can add 'freshness' and flavor to wines
####  4 - residual sugar: the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with  greater than 45 grams/liter are considered sweet
####  5 - chlorides: the amount of salt in the wine
####  6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine
#### 7 - total sulfur dioxide: amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
#### 8 - density: the density of water is close to that of water depending on the percent alcohol and sugar content
#### 9 - pH: describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
#### 10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
#### 11 - alcohol: the percent alcohol content of the wine
####12 - quality (score between 0 and 10)


### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
#### the rating feature concluded from the quality variabl.
### Did you create any new variables from existing variables in the dataset?
#### yes, the 

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
#### There were no unusual distibutions among all the variables, all the attributes are bormally distibuted except for the alcohol variable, it gad several peaks 

# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
## Correlation between Variables:
 #### by removing the X and rating columns 

correlation_matrix <- cor(
  ww %>%
    dplyr::select(-X) %>%
    dplyr::select(-rating) %>%
    mutate(
      quality = as.numeric(quality)
    )
  )

## Results:
View(correlation_matrix)
ggplot(aes(x=density,y=alcohol), data=ww)+
  geom_point(alpha=1/10)
```
```{r echo=FALSE, Bivariate_Plots}

ggplot(aes(x=chlorides,y=alcohol), data=ww)+
  geom_point(alpha=1/10)
```

```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=alcohol,y=quality), data=ww)+
  geom_point(alpha=1/10)

names(ww)
```


```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=density,y=quality), data=ww)+
  geom_point(alpha=1/20)
```




```{r echo=FALSE, Univariate_Plots}
ggplot(aes(x=residual.sugar,y=density), data=ww)+
  geom_point(alpha=1/10)
```


```{r echo=FALSE, Univariate_Plots}
qplot(sulphates, data=ww)
```

```{r echo=FALSE, Univariate_Plots}
qplot(sulphates, data=ww)
```













# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \

Regarding the quality, Alcohol, shows a high direct correlation with it with  a value of 0.4.
then comes the density and the quality amount, with high negative correlation 


the highest correlation in general happens to be between:
1- Density and Alcohol, the more you add the alcohol, the less dense the wine becomes (-0.78)
2- Density and Residual Sugar, the more the sugar is added the denser the wine gets (0.83)
3- Also there is a correlation between Total Sulfur Dioxide and Residual Sugar, Density and Free Sulfur Dioxide with the values (0.4, 0.5298, 0.615) respectively 



### Did you observe any interesting relationships between the other features \
there is a correlation between Total Sulfur Dioxide and Residual Sugar, Density and Free Sulfur Dioxide with the values (0.4, 0.5298, 0.615) respectively 

### What was the strongest relationship you found?
Density and Residual Sugar, the more the sugar is added the denser the wine gets (0.83)

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

## Adding the quality to th previous Alcohol vs Density plot:
### There is no evidence that the density has nothing to do with the quality, but the alchol has a strong correlation value with the quality as per the plot below.
ggplot(aes(x = alcohol, y = density), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')


```

```{r echo=FALSE, Multivariate_Plots}

## Adding the quality to th previous Alcohol vs Cholorides plot:
### There is no evidence that the Chlorides has nothing to do with the quality, but the alchol has a strong correlation value with the quality as per the plot below.

  
  
ggplot(aes(x = alcohol, y = chlorides), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')

```

```{r echo=FALSE, Multivariate_Plots}

## Adding the quality to th previous Residual Sugar vs Density plot:
### 
ggplot(aes(x = residual.sugar, y = density), data = subset(ww, residual.sugar < 30)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')


```


```{r echo=FALSE, Multivariate_Plots}

## Sulphats vs alcohol vs Quality
### Wines with higher levels of sulphates and alchol are more likely t have higher quality.
ggplot(aes(x = alcohol, y = sulphates), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')


```
```{r echo=FALSE, Multivariate_Plots}

## Acids vs Quality

### no Variance noted between the Volatile acids, citric acids vs the quality 
ggplot(aes(x = citric.acid, y = volatile.acidity), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```
```{r echo=FALSE, Multivariate_Plots}

## Acids vs Quality
### Citric acids vs the fixed acidity.
### no Variance noted between the Volatile acids, citric acids vs the quality 
ggplot(aes(x = citric.acid, y = fixed.acidity), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```
```{r echo=FALSE, Multivariate_Plots}

## Sulfur dioxide  vs Quality

### Both oxides are dependent on each other but quality does not depend on any of them.
ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
facet_wrap(~rating) 

```
```{r echo=FALSE, Multivariate_Plots}

##Sulphates vs Density  vs Quality

### Both oxides are dependent on each other but quality does not depend on any of them.
ggplot(aes(x = sulphates, y = density), data = subset(ww)) + 
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = 'seq')+
  facet_wrap(~rating) 

```
  
# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------


> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!